\documentclass{article}\usepackage{amsmath}\begin{document}\title{Simplex Solver}\maketitle\begin{flushleft}\textbf{Problem}\end{flushleft}\begin{flushleft}Given the following linear system and objective function, find the optimal solution.\end{flushleft}\begin{equation*}\max{ 3.0x_1 + 2.0x_2 } \\ \end{equation*}\[\left\{\begin{array}{c} x_1 + 2.0x_2 \leq 8.0 \\  4.0x_1 \leq 16.0 \\  4.0x_2 \leq 12.0 \\ \end{array}\right.\]\begin{flushleft}\textbf{Solution}\end{flushleft}\begin{flushleft}Add slack variables to turn all inequalities to equalities.\end{flushleft}\[\left\{\begin{array}{c} x_1 + 2.0x_2 + s_1 = 8.0 \\  4.0x_1 + s_2 = 16.0 \\  4.0x_2 + s_3 = 12.0 \\ \end{array}\right.\]\begin{flushleft}Create the initial tableau of the new linear system.\end{flushleft}\begin{equation*}\begin{bmatrix}\begin{array}{ccccc|c}x_1 &x_2 &s_1 &s_2 &s_3 &b \\ \hline1.0 & 2.0 & 1 & 0 & 0 & 8.0 \\4.0 & 0.0 & 0 & 1 & 0 & 16.0 \\0.0 & 4.0 & 0 & 0 & 1 & 12.0 \\ \hline-3.0 & -2.0 & 0 & 0 & 0 & 0 \\\end{array}\end{bmatrix}\begin{array}{c}\\s_1 \\s_2 \\s_3 \\\\\end{array}\end{equation*}\begin{flushleft}There are negative elements in the bottom row, so the current solution is not optimal. Thus, pivot to improve the current solution. The entering variable is $x_1$ and the departing variable is $s_2$.\end{flushleft}\begin{flushleft}Perform elementary row operations until the pivot element is 1 and all other elements in the entering column are 0.\end{flushleft}\begin{equation*}\begin{bmatrix}\begin{array}{ccccc|c}x_1 &x_2 &s_1 &s_2 &s_3 &b \\ \hline0.0 & 2.0 & 1.0 & -0.25 & 0.0 & 4.0 \\1.0 & 0.0 & 0.0 & 0.25 & 0.0 & 4.0 \\0.0 & 4.0 & 0.0 & 0.0 & 1.0 & 12.0 \\ \hline0.0 & -2.0 & 0.0 & 0.75 & 0.0 & 12.0 \\\end{array}\end{bmatrix}\begin{array}{c}\\s_1 \\x_1 \\s_3 \\\\\end{array}\end{equation*}\begin{flushleft}There are negative elements in the bottom row, so the current solution is not optimal. Thus, pivot to improve the current solution. The entering variable is $x_2$ and the departing variable is $s_1$.\end{flushleft}\begin{flushleft}Perform elementary row operations until the pivot element is 1 and all other elements in the entering column are 0.\end{flushleft}\begin{equation*}\begin{bmatrix}\begin{array}{ccccc|c}x_1 &x_2 &s_1 &s_2 &s_3 &b \\ \hline0.0 & 1.0 & 0.5 & -0.125 & 0.0 & 2.0 \\1.0 & 0.0 & 0.0 & 0.25 & 0.0 & 4.0 \\0.0 & 0.0 & -2.0 & 0.5 & 1.0 & 4.0 \\ \hline0.0 & 0.0 & 1.0 & 0.5 & 0.0 & 16.0 \\\end{array}\end{bmatrix}\begin{array}{c}\\x_2 \\x_1 \\s_3 \\\\\end{array}\end{equation*}\begin{flushleft}There are no negative elements in the bottom row, so we know the solution is optimal. Thus, the solution is: \end{flushleft}\begin{equation*}s_1 = 0, s_2 = 0, s_3 = 4.0, x_1 = 4.0, x_2 = 2.0, z = 16.0\end{equation*}\end{document}